Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-15 IEEE_ConferenceOnGames/new_article.tex
more work

    \section{State of the art}
    \label{SoA}
    In this Section we present PCG works relevant to the Level generation for the
    Angry Birds video game. Renz et al. \cite{Renz, Renz2015TheAB} organized an AI
    competition with the aim of developing intelligent agents that could play new
    Angry Birds levels better than the best human players. The competition was then
    collocated in other AI conferences. In order to train and test these agents,
    more novel and some times difficult levels must be generated. A related
    competition was then proposed by Stephenson et al. \cite{Stephenson} this time
    with the goal of developing a level generator for the game. Several generators
    have been proposed in the following years of competition. M. Kaidan et al.
    \cite{Kaidan2015} propose a model in which the generation of content is
    controlled by the skill of a player while actively playing the game, the moment
    the player finishes a certain level the algorithm obtains the score and
    generates the next level to be played. Y. Jiang et al. \cite{Jiang2017} proposed
    the use of predefined letter style patterns and combined to create a small pool
    of words and letters used to be combined in order to create levels with text on
    them, the participants would play the game and after that they would be asked if
    they liked the layout of the levels. A search based approach has been also used
    in several works. Laura Calle \cite{lauracalle} proposed a method in which the 
    pieces in the game are arranged in groups using the gravity of the game to 
    create structures that are later evolved into more complex levels, this method 
    was used as a base to work with since it used a modified version of the angry 
    birds game that was able to obtain the data of a level after a simulation.
    %%%
    % TO DO: Laura´s work here 
    %%%

@@

    %Second, after removing a percentage according to the missing pieces an individual
    %remaining pieces are checked for their positions in comparison to the original
    %ones before the simulation, using the formula 1 each piece within the individual
    %is evaluated according to the center point of the piece, if the piece moved more
    %than a certain threshold half the percentage for one piece is removed from the
    %remaining percentage, after this using formula 2 the position of the piece
    %is further checked, this time by the difference in angle from the original one,
    %in case the difference of angles is more than a certain threshold half a piece
    %value is further removed from the total.

    %\begin{equation}
    %    \begin{aligned}
    %    error_{xy} = \left\{ \,
    %        \begin{IEEEeqnarraybox}[][c]{l?s}
    %            \IEEEstrut
    %            0.1 & if $ 0.08 > d $, \\
    %            \frac{100}{lenght\_pieces} * 0.5 & if $ 0.08 < d $.
    %            \IEEEstrut
    %        \end{IEEEeqnarraybox}
    %    \right. \\
    %    where: d=\sqrt{(x_2 - x_1)^2 + (y_2 - y_1)^2}    
    %    \end{aligned}
    %    \label{error_xy}
    %\end{equation}

   % \begin{equation}
   %     \begin{aligned}
   %     error_{r} = \left\{ \,
   %         \begin{IEEEeqnarraybox}[][c]{l?s}
   %             \IEEEstrut
   %             0 & if $ -5 \le r \le 5 $, \\
   %             \frac{100}{lenght\_pieces} * 0.5 & if $ r < 5 \: or \: r > 5 $.
   %             \IEEEstrut
   %         \end{IEEEeqnarraybox}
   %     \right. \\
   %     where: r= |rotation_0| - |rotation_f|   
   %     \end{aligned}
   %     \label{error_angle}
   % \end{equation}

  % Finally, the total height of the remaining group of pieces is calculated by
  %  combining the bonding boxes of all pieces, then the individuals are ordered by
  %  the obtained percentage and then further ordered by the total height, this way
  %  the most balanced ones will have the best probability to be selected together
  %  for a crossover operation to create the next generation.

 %   \begin{figure}[htbp]
 %   \centerline{\includegraphics[width=80mm]{Images/result_example.png}}
 %   \caption{Result samples of current system.}
 %   \label{results_old}
 %   \end{figure}

 %   Since there is a need to create complex structures than can cover a great
 %   section of the play map, a method to evaluate the distribution is needed, for
 %   this instance we propose using the rule of thirds \cite{DarrenRowse} which is a
 %   guideline used while creating images, the purpose is to create
 %   aesthetical pleasant images where the focus point or the most important area is
 %   at the center of the image as shown on figure \ref{rule_of_thirds}, using this
 %   rule different distribution groups or masks are created has shown on figure
 %   \ref{rule_of_thirds_masks} in order to use them before the simulation of an
 %   individual by placing the pieces in order to distribute them in a way that can cover
 %   more area and create more balanced levels instead of using a tower like
 %   distribution as shown on figure \ref{results_old}.

 %   \begin{figure}[htbp]
 %       \centerline{\includegraphics[width=80mm]{Images/ruleofthirds_example.png}}
 %       \caption{Rule of thirds example.}
 %       \label{rule_of_thirds}
 %   \end{figure}

 %   \begin{figure}[htbp]
 %       \centerline{\includegraphics[width=80mm]{Images/mask_distribution.png}}
 %       \caption{Masks created using the rule of thirds.}
 %       \label{rule_of_thirds_masks}
 %   \end{figure}

    Using this same distribution a new evaluation will be added in which according
    to how much of the area is filled a score will be given and this score will be
    used as a new sorting value at the moment of selecting individuals for the
    crossover operation, since as shown on figure \ref{rule_of_thirds_masks} there
    is a possibility that a huge mask is given to an individual it can be possible
    that the amount of pieces will not be able to cover the full area given, so in
    order to prevent this kind of problem we also proposed to give each mask a
    value that refers to the minimum amount of pieces and height required to fill
    the most of the mask, then check the amount of pieces and height before
    simulation for a single individual and then randomize which mask of the ones
    that can be used will be assigned to that individual, this way the results shown
    on figure \ref{rule_of_thirds_result} are obtained, this however gave us a new
    problem because as shown in the figure and on figure \ref{rule_of_thirds_applied} as well, the were
    instances in which a certain combination of pieces could be to big for the area
    and this pieces resulted in a level that had structural flaws from the beginning
    which cold provoke the entire structure and immeadiately nearby others to fall
    at the beginning of the simulation however this allowed us to realize that predefined 
    composites could provide problems to the generated structures therefore this approach 
    was not the best for this problem.

    \begin{figure}[htbp]
        \centerline{\includegraphics[width=80mm]{Images/chromosome_thirds.png}}
        \caption{Rule of thirds applied to an individual.}
        \label{rule_of_thirds_applied}
    \end{figure}

@@

    \subsection{New approach}

    The first differece that the new approach has is that the masks that are assigned
    to each individual are no longer pre-generated ones, this is because we decided
    that pre-generating them would not allow the generator to \textit{evolve}
    because the individuals would be limited to a certain number of possibles
    combinations which will in turn prevent diversity in the population, so in order
    to prevent this the modification to this section allows the individuals to
    generate a mask in which a number of columns is defines by default but the
    amount of pieces in each one is random, however the total of assigned pieces in
    all columns must be the same as the total pieces allowed in each individual
    which is a number assigned before beginning the system, an example of this
    can be seen in figure \ref{new_chrom} where the total of pieces in a individual
    is 18 and this pieces are distributed on all the columns available. This is
    further explained in this section.

    The competition also requires that a generator follows a certain ruleset this is 
    emulated by providing a file to the system to simulate the ruleset provided at the 
    competition, this \textit{text file
    (parameters.txt)} provides the specifications that must be met in the
    competition, the parameters in the file are as follows:

    \begin{itemize}
        \item Total required levels.
        \item Restrictions to pieces or piece-material combinations that cannot be
        placed or used in the levels, in this case if the file specifies that a
        \textit{CIRCLE} made out of \textit{ICE} cannot be placed, so the system must be
        able to verify that said combination is not used.
        \item Total number of pigs, this value can be a single value or a range of
        values that represent how many pigs must be placed on a level.
        \item The amount of time in which the system must give an output.
    \end{itemize}

@@

    Using this as a base a new system diagram is proposed as shown of figure
    \ref{new_model} in which the file with parameters is integrated with the input
    data  the system requires, the way the individuals of the population are further
    presresented and the way the simulation phase is taken, the method to
    representate the individuals is proposed by dividing the individual on three
    layers represented as follows:

    \begin{itemize}
        \item Element layer: also known as pieces layer, this layer represents the
        elements assigned to a particular individual by an array of numbers from 0
        to infinite, where each number represents a reference to the class of a piece
        in the selection pool including the ones generated by the OOE algorithm, the
        representetion of this is was shown in figure \ref{old_chrom}.
        \item  Mask layer: this layer is represented by a 7 digit array in which the
        value assigned in each column represents the amount of pieces from the
        element layer that will be assigned to said column, a representation of this
        is shown on figure \ref{mask_layer}.
        \item Enemy layer: also known as pig layer, this layer is represented by two
        elements a single numeric value representing the total amount of pigs
        assigned to the level and an array of locations for the placement of this
        same pigs, a representation of it is shown on figure \ref{enemy_layer}. 
    \end{itemize}

@@

    In order to select the parents of the next generation the parents are ordered by
    their fitness, then the system can choose between 2 options for the selection,
    tournament selection or roulette selection:

    \begin{itemize}
        \item Tournament selection: in order to select the parent by this method the
        system checks the amount of crossovers assigned, then calculates the ammount
        of parent required to compete in pairs, then by doing a descendent selection
        the slots for the tournament are filled and the winners are calculated in
        each group.
        \item Roulette selection: this method of selection was first decided to be
        used directly by giving a proportional amount of chances to be selected
        according to the fitness of a individual, however since one of the
        objectives is to obtain variety of levels the roulette must first be ranked
        in order to ensure that low fitness individuals can be selected as parents.
    \end{itemize}

>>>>

    \section{State of the art}
    \label{SoA}
    In this Section we present PCG works relevant to the Level generation for the
    Angry Birds video game. Renz et al. \cite{Renz, Renz2015TheAB} organized an AI
    competition with the aim of developing intelligent agents that could play new
    Angry Birds levels better than the best human players. The competition was then
    collocated in other AI conferences. In order to train and test these agents,
    more novel and some times difficult levels must be generated. A related
    competition was then proposed by Stephenson et al. \cite{Stephenson} this time
    with the goal of developing a level generator for the game. Several generators
    have been proposed in the following years of competition. M. Kaidan et al.
    \cite{Kaidan2015} propose a model in which the generation of content is
    controlled by the skill of a player while actively playing the game, the moment
    the player finishes a certain level the algorithm obtains the score and
    generates the next level to be played. Y. Jiang et al. \cite{Jiang2017} proposed
    the use of predefined letter style patterns and combined to create a small pool
    of words and letters used to be combined in order to create levels with text on
    them, the participants would play the game and after that they would be asked if
    they liked the layout of the levels. A search based approach has been also used
    in several works. 
    %%%
    % TO DO: Laura´s work here 
    %%%

@@

    \subsection{Structure Positioning}

     The positioning of structures in the map is important in order to generate
     levels that are fun and aesthetic. For this we need a method to indcate the
     positioning of each structure. We propose the use of a mask where the
     positioning of each pile and the number of blocks it contains is specefied.
     Again these masks could be manually created, for instance using the rule of
     thirds \cite{DarrenRowse} which is a guideline used while creating images,
     the purpose is to create aesthetical pleasant images where the focus point
     or the most important area is at the center of the image as shown on figure
     \ref{rule_of_thirds}, using this rule different distribution groups or
     masks are created has shown on figure \ref{rule_of_thirds_masks} in order
     to use them before the simulation of an individual by placing the pieces in
     order to distribute them in a way that can cover more area and create more
     balanced levels instead of using a tower like distribution as shown on
     figure \ref{results_old}. There is an option in the system to generate in a
     first step a collection of masks, following this rules, but this qpproach
     has a drawback. Pre-generating masks would not allow the generator to
     \textit{evolve} because the individuals would be limited to a certain
     number of possibles combinations which will in turn prevent diversity in
     the population, so in order to prevent this we proposed an approach that
     allows the individuals to generate a mask in which a number of columns is
     defined by default but the amount of pieces in each one is random, however
     the total amount of assigned blockes in all columns must be the same as the
     total pieces allowed in each individual. The number of maximum blocks is as
     a parameter. An example of this can be seen in figure \ref{new_chrom} where
     the total of pieces in a individual is 18 and this pieces are distributed
     on all the columns available.

    \begin{figure}[htbp]
        \centerline{\includegraphics[width=80mm]{Images/ruleofthirds_example.png}}
        \caption{Rule of thirds example.}
        \label{rule_of_thirds}
    \end{figure}

    \begin{figure}[htbp]
        \centerline{\includegraphics[width=80mm]{Images/mask_distribution.png}}
        \caption{Masks created using the rule of thirds.}
        \label{rule_of_thirds_masks}
    \end{figure}

@@

   % The competition also requires that a generator follows a certain ruleset this is 
   % emulated by providing a file to the system to simulate the ruleset provided at the 
   % competition, this \textit{text file
   % (parameters.txt)} provides the specifications that must be met in the
   % competition, the parameters in the file are as follows:

   % \begin{itemize}
   %     \item Total required levels.
   %     \item Restrictions to pieces or piece-material combinations that cannot be
   %     placed or used in the levels, in this case if the file specifies that a
   %     \textit{CIRCLE} made out of \textit{ICE} cannot be placed, so the system must be
   %     able to verify that said combination is not used.
   %     \item Total number of pigs, this value can be a single value or a range of
   %     values that represent how many pigs must be placed on a level.
   %     \item The amount of time in which the system must give an output.
   % \end{itemize}

@@

    A diagram showng the main components and data flow of the proposed system
    is shown in Figure \ref{new_model} in which the parameter file of the competition 
    is integrated as an input. The layers of the system is exaplained next:

    \begin{itemize}
        \item Composite Generator:  This layer represents the composite blocks
        assigned to a particular individual as list of integer numbers, where
        each number represents a reference to a composite object including the
        ones generated by the OOE algorithm, the representetion of this is was
        shown in figure \ref{old_chrom}.
        \item Structure layer: Individuals are represented by a list of seven
        integers, representing a mask in which integers assigned in each column
        represents the amount of composite blocks from the element layer that
        will be assigned to said column, a representation of this is shown on
        figure \ref{mask_layer}.
        \item Enemy layer: also known as pig layer, this layer is represented by
        two elements a single numeric value representing the total amount of
        pigs assigned to the level and an array of locations for the placement
        of this same pigs, a representation of it is shown on figure
        \ref{enemy_layer}. 
    \end{itemize}

@@

    \subsection{Evolution of Levels}

    In order to select the parents of the next generation the parents are ordered by
    their fitness. There are two types of selection:

    \begin{itemize}
    % This needs more work.    
        \item Tournament selection: in order to select the parent by this method the
        system checks the amount of crossovers assigned, then calculates the ammount
        of parent required to compete in pairs, then by doing a descendent selection
        the slots for the tournament are filled and the winners are calculated in
        each group.
        \item Roulette selection: this method of selection was first decided to be
        used directly by giving a proportional amount of chances to be selected
        according to the fitness of a individual, however since one of the
        objectives is to obtain variety of levels the roulette must first be ranked
        in order to ensure that low fitness individuals can be selected as parents.
    \end{itemize}
